---
title: "Thompson Sampling Ad Auction"
author: "Jared Brooks"
date: "9/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Skills Practice

This is for practicing some DS skills. I am following a post from Count Bayesie here: https://www.countbayesie.com/blog/2020/9/26/learn-thompson-sampling-by-building-an-ad-auction

## Data Setup

First I'm just setting up example data to play with

```{r}
ad_info <- data.frame("advertiser" = c("buymystuff.com", "eyespyonyou.com", "commodityfetishismllc.com"),
                      "bid" = c(0.7, 0.5, 0.4), "views" = c(2000, 0,0), 
                      "clicks" = c(47,0,0), "ctr" = c(0.025, 0.012, 0.045))
library(knitr)
kable(ad_info)
```

Given the previous views and clicks for "buymystuff.com" we can look at a likelihood estimate plot for its click-through-rate

```{r}
library(ggplot2)
ggplot() + stat_function(fun = function(x) dbeta(x, 47,1953), color="black") + xlim(0, 0.04)
```

Now we pretend like we don't know the CTRs for the other two ad bidders, but that we do have a lot of information about how campaigns perform in general. We generate this using a beta distribution assuming an average CTR of 0.02 from 500 campaigns.

```{r}
past_ctrs <- rbeta(500, 2, 98, ncp = 0)

prior_est <- MASS::fitdistr(past_ctrs, dbeta,
#choosing these shape parameters because our mean(past_ctrs) is about 0.02                    
                            start = list(shape1 = 2, 
                                         shape2 = 98))
alpha_prior <- prior_est$estimate[1]
beta_prior <- prior_est$estimate[2]

ggplot() + geom_histogram(aes(past_ctrs), binwidth = .002) + stat_function(fun = function(x) dbeta(x, alpha_prior, beta_prior), color = "red",size = 1) + xlab("CTR")
```

Now, given the CTR from one bidder and using a beta distribution on the previous campaigns to estimate the CTRs from the other two bidders, and including their bids, we can plot the distributions of expected values

```{r}
ggplot() + stat_function(fun = function(x) dbeta(x/ad_info$bid[1], ad_info$clicks[1],(ad_info$views[1]-ad_info$clicks[1])), color="red", size=1) + stat_function(fun = function(x) dbeta(x/ad_info$bid[2], alpha_prior, beta_prior), color="green", size=1) + stat_function(fun = function(x) dbeta(x/ad_info$bid[3], alpha_prior, beta_prior), color="blue", size=1) + xlim(0,0.03) + ggtitle("Distribution of Expected Values")
```

## Sampling Frequency for who wins the bids

We now have enough to make estimates of the expected values to see who will win the bids by sampling from our distributions.

```{r}
#number of samples we want
N <- 10000
#BuyMyStuff.com
ad_A <- ad_info$bid[1]*rbeta(N,
                             ad_info$clicks[1] + alpha_prior,
                             (ad_info$views[1]-ad_info$clicks[1]) +
                             beta_prior)

#EyeSpyOnYou.com
ad_B <- ad_info$bid[2]*rbeta(N, 
                             ad_info$clicks[2] + alpha_prior,
                             (ad_info$views[2]-ad_info$clicks[2]) + 
                             beta_prior)

#CommodityFetishismLLC.com
ad_C <- ad_info$bid[3]*rbeta(N, 
                             ad_info$clicks[3] + alpha_prior,
                             (ad_info$views[3]-ad_info$clicks[3]) +
                             beta_prior)
```

Let's compare the sampling frequency of A and B

```{r}
ggplot() + geom_histogram(aes(ad_A-ad_B), bins = 12) + geom_vline(xintercept = 0) + ggtitle("Sampling Difference for EV_A and EV_B")
```

Let's compare B and C against A to see their probably of getting picked over A

```{r}
ad_info$p_gt_A <- c(0.5, sum(ad_A < ad_B)/N, sum(ad_A < ad_C)/N)
```

Ad A gets 0.5 as it is being compared against itself.

Now lets normalize these to get weighted probabilities that add to 1

```{r}
ad_info$weighted_p_gt_A <- ad_info$p_gt_A/(sum(ad_info$p_gt_A))
```

Now we sample from these calculated probabilites to see how many new views each ad gets, and use our "known" CTRs to estimate the new clicks for each add

```{r, message=FALSE}
library(dplyr)
winners_df <- sample(ad_info$advertiser,
                     prob = ad_info$weighted_p_gt_A,
                     replace = TRUE,
                     size=500) %>%  
  table() %>%  
  as.data.frame()

ad_info$new_views <- winners_df$Freq
ad_info$new_clicks <- rbinom(c(1,1,1),ad_info$new_views,ad_info$ctr)
```

Now lets look at a table of the results

```{r}
kable(ad_info[,c(1,2,3,4,8,9)])
```

And we can see the updated distributions

```{r}
ggplot() + 
  stat_function(fun = function(x) dbeta(x/ad_info$bid[1], ad_info$clicks[1] + ad_info$new_clicks[1], (ad_info$views[1]-ad_info$clicks[1]) + (ad_info$new_views[1]-ad_info$new_clicks[1])), color="red", size=1) + 
  
  stat_function(fun = function(x) dbeta(x/ad_info$bid[2], alpha_prior + ad_info$new_clicks[2], beta_prior + (ad_info$new_views[2]-ad_info$new_clicks[2])), color="green", size=1) + 
  
  stat_function(fun = function(x) dbeta(x/ad_info$bid[3], alpha_prior + ad_info$new_clicks[3], beta_prior + (ad_info$new_views[3]-ad_info$new_clicks[3])), color="blue", size=1) + xlim(0,0.03) + 
  
  ggtitle("Distribution of Expected Values")

```

Now we can use this to calculate and look at new sampling probabilities

```{r}
#number of samples we want
N <- 10000
#BuyMyStuff.com
ad_A <- ad_info$bid[1]*rbeta(N,
                             ad_info$clicks[1] + ad_info$new_clicks[1] + alpha_prior,
                             (ad_info$views[1]-ad_info$clicks[1]) + (ad_info$new_views[1]-ad_info$new_clicks[1]) +
                             beta_prior)

#EyeSpyOnYou.com
ad_B <- ad_info$bid[2]*rbeta(N, 
                             ad_info$clicks[2] + ad_info$new_clicks[2] + alpha_prior,
                             (ad_info$views[2]-ad_info$clicks[2]) + (ad_info$new_views[2]-ad_info$new_clicks[2]) +
                             beta_prior)

#CommodityFetishismLLC.com
ad_C <- ad_info$bid[3]*rbeta(N, 
                             ad_info$clicks[3] + ad_info$new_clicks[3] + alpha_prior,
                             (ad_info$views[3]-ad_info$clicks[3]) + (ad_info$new_views[3]-ad_info$new_clicks[3]) +
                             beta_prior)

ad_info$p_gt_A <- c(0.5, sum(ad_A < ad_B)/N, sum(ad_A < ad_C)/N)
ad_info$weighted_p_gt_A <- ad_info$p_gt_A/(sum(ad_info$p_gt_A))

kable(ad_info[,c(1,6,7)])
```


## Continuous updating


